Automated Lensing Learner - I: An automated Strong Lensing Identification Pipeline

### 2017-03-23 ms.tex
Went through matt's new text

%-------------------------------------------------%
\begin{abstract} 
Gravitational lensing offers a direct probe of the underlying mass
distribution of lensing systems, a window to the high redshift
universe, and a geometric probe of cosmological models.  The advent of
large scale surveys such as the Large Synoptic Sky Telescope and
Euclid has prompted a need for automatic and efficient identification
of strong lensing systems.  We present {\em (ALL) Automated Lensing
  Learner}, a strong lensing identification pipeline that will be
publicly released as open source software.  In this first application,
we employ a fast feature extraction method, Histogram of Oriented
Gradients (HOG), to capture edge patterns that are characteristic of
strong gravitational arcs in galaxy-galaxy lensing.  We use logistic
regression to train a supervised classifier model on the HOG of HST-
and LSST-like images.  We use an area under the curve (AUC) of a
Receiver Operating Characteristic curve to assess model performance;
$AUC=1.0$ is an ideal trained classifier, and $AUC=0.5$ does not
provide any classification information.  Our best performing models on
a training set of 10,000 lens containing images and 10,000 non-lens
containing images exhibit an AUC of 0.975 for an HST-like sample.
However, for one orbit of LSST, our model only reaches an AUC of
0.625.  For 10-year mock LSST observations, the AUC improves to 0.809.
Model performance appears to continually improve with the size of the
training set.  Models trained on fewer images perform better in
absence of the lens galaxy, which can decrease the false positive rate
by blending with light from a lensed galaxy image near the central
lensing galaxy.  However, with larger training data sets, information
from the lensing galaxy improves model performance.  Our
results demonstrate efficient and effective methods for automated strong
lensing identification. The simple classifier used here easily
scales to large data sets on a computing cluster and is easily
reproduced on a personal laptop, using existing open source tools
in both cases.
\end{abstract}
%-------------------------------------------------%

@@

% Benefits of our method
We emphasize that simple linear classifiers, such as
logistic regression, are scalable and relatively easy to parallelize
with open source tools such as {\em
Spark}\footnote{http://spark.apache.org/}.  Initial checks
indicate that HOG feature extraction plus a linear
classifier captures much of the complexity in the arc finding problem.
These results indicate that effective image classification of
galaxy-galaxy strong lensing systems can be achieved without
large scale computing resources, or specialized hardware like GPUs,
but can be scaled to large datasets on a computing cluster if needed.

@@

% Interpreting TP and FP rates
Finally, we must take the class balance, or relative number of lens to
non-lens systems, into account when assessing a metric.  The
Purity-Recall (purity-completeness) metric is sensitive to the ratio
of lens and non-lens systems in the data.  Both of our training and
test data sets have $50\%$ lensed and $50\%$ unlensed images, which is
not expected in observations.  The ROC curve metric is insensitive to
the ratio, but is sensitive to the sampling.  Given alternative lens
and non-lens sample splittings, our true positive and false positive
rates in the ROC curves would stay the same, making the ROC curve a
more standard metric in the literature.  On the other hand, the ROC
curves show a representative rate for lens and source distributions
that are evenly sampled.  We do {\it not} expect this sampling to be
representative of what we might expect from a an observational survey.

\acknowledgments CA acknowledges support from the Enrico Fermi
Institute at the University of Chicago, and the University of Chicago
Provost's Office. This work was also supported in part by the Kavli
Institute for Cosmological Physics at the University of Chicago
through grant NSF PHY-1125897 and an endowment from the Kavli
Foundation and its founder Fred Kavli.  NL would like to thank the
funding support from NSFC, grant no.11503064, and Shanghai Natural
Science Foundation, grant no. 15ZR1446700.
\lastpagefootnotes

>>>>

%-------------------------------------------------%
\begin{abstract} 
Gravitational lensing offers a direct probe of the underlying mass
distribution of lensing systems, a window to the high redshift
universe, and a geometric probe of cosmological models.  The advent of
large scale surveys such as the Large Synoptic Sky Telescope and
Euclid has prompted a need for automatic and efficient identification
of strong lensing systems.  We present (1) {\em (ALL) Automated
  Lensing Learner}, a strong lensing identification pipeline that will
be publicly released as open source software, and (2) a publicly
available mock LSST dataset with strong galaxy-galaxy lenses.  In this
first application of the pipeline, we employ a fast feature extraction
method, Histogram of Oriented Gradients (HOG), to capture edge
patterns that are characteristic of strong gravitational arcs in
galaxy-galaxy lensing.  We use logistic regression to train a
supervised classifier model on the HOG of HST- and LSST-like images.
We use an area under the curve (AUC) of a Receiver Operating
Characteristic curve to assess model performance; $AUC=1.0$ is an
ideal trained classifier, and $AUC=0.5$ does not provide any
classification information.  Our best performing models on a training
set of 10,000 lens containing images and 10,000 non-lens containing
images exhibit an AUC of 0.975 for an HST-like sample.  However, for
one orbit of LSST, our model only reaches an AUC of 0.625.  For
10-year mock LSST observations, the AUC improves to 0.809.  Model
performance appears to continually improve with the size of the
training set.  Models trained on fewer images perform better in
absence of the lens galaxy, the removal of which can decrease the
false positive rate due to the blending of light between source images
and the lens galaxy.  However, with larger training data sets,
information from the lens galaxy improves model performance.  Our
results demonstrate efficient and effective methods for automated
strong lensing identification. With existing open source tools, this
example linear classifier both runs on a personal laptop and can
easily scale to large data sets on a computing cluster.
\end{abstract}
%-------------------------------------------------%

@@

% Benefits of our method
We emphasize that simple linear classifiers, such as logistic
regression, are scalable and relatively easy to parallelize with open
source tools such as {\em Spark}\footnote{http://spark.apache.org/}.
Our work indicates that HOG feature extraction plus a linear
classifier captures much of the complexity in the arc finding problem.
{\it These results indicate that effective image classification of
  galaxy-galaxy strong lensing systems can be achieved without large
  scale computing resources, or specialized hardware like GPUs.}
These methods also scale for large datasets on a computing cluster, if
needed.

@@

% Interpreting TP and FP rates
Finally, we must take the class balance, or relative number of lens to
non-lens systems, into account when assessing a metric.  The
Purity-Recall (purity-completeness) metric is sensitive to the ratio
of lens and non-lens systems in the data.  Both of our training and
test data sets have $50\%$ lensed and $50\%$ unlensed images, which is
not expected in observations.  The ROC curve metric is insensitive to
the ratio, but is sensitive to the sampling.  Given alternative lens
and non-lens sample splittings, our true positive and false positive
rates in the ROC curves would stay the same, making the ROC curve a
more standard metric in the literature.  On the other hand, the ROC
curves show a representative rate for lens and source distributions
that are evenly sampled.  We do {\it not} expect this sampling to be
representative of what we might expect from a an observational survey.
We leave these additional challenges to future work.

\acknowledgments CA acknowledges support from the Enrico Fermi
Institute at the University of Chicago, and the University of Chicago
Provost's Office. NL would like to thank the funding support from
NSFC, grant no.11503064, and Shanghai Natural Science Foundation,
grant no. 15ZR1446700. This work was also supported in part by the
Kavli Institute for Cosmological Physics at the University of Chicago
through grant NSF PHY-1125897 and an endowment from the Kavli
Foundation and its founder Fred Kavli.emacs -nw 
\lastpagefootnotes
